In quantum mechanics, variational principle is usually used to obtain ground state properties of physical systems.
We first assume a variational wave function $\Phi_t(\alpha,\vec{R})$ with parameters $\alpha$.
This wave function should be as close as to exact one for better approximation.
The energy of this trail wave function can be calculated and minimized with regard to variational parameters
\begin{equation}\label{var}
	<H(\alpha)>=\frac{\int d\vec{R}\Phi_t^\ast(\alpha,\vec{R})\hat{H}(\vec{R})\Phi_t(\alpha,\vec{R})}{\int d\vec{R}\Phi_t^\ast(\alpha,\vec{R})\Phi_t(\alpha,\vec{R})};\ \frac{dH}{d\alpha}=0.
\end{equation}
This lowest energy of the trail wave function is an approximation of exact ground energy of the system.

Monte Carlo(MC) method is best applicable for high demential integrals as usually seen in quantum many-body problems.
Combining variational principle and MC method, we can solve integrals in Eq. \ref{var}.
In order to evaluate this integral, we define two qualities local energy and probability
\begin{equation}\label{local}
		E_L(\vec{R},\alpha)=\frac{\hat{H}\Phi_t(\alpha,\vec{R})}{\Phi_t(\alpha,\vec{R})}
\end{equation}
\begin{equation}\label{local}
		P(\vec{R},\alpha)=\frac{\Phi_t^\ast(\alpha,\vec{R})\Phi_t(\alpha,\vec{R})}{\int d\vec{R}\Phi_t^\ast(\alpha,\vec{R})\Phi_t(\alpha,\vec{R})}.
\end{equation}
Having above two equations, Eq. \ref{var} can be expressed as 
\begin{equation}
	<H(\alpha)>=\int d\vec{R}P\cdot E_L \approx \frac{1}{N}\sum_{i=1}^{N-1}P\cdot E_L
\end{equation}
which can be evaluated through MC simulations.

The report is organized as follows. 
In Sec. \ref{phys}, we give an introduction to the physical system we studied in detail.
In Sec. \ref{method}, we show both famous Metropolis algorithm and our algorithm based on it.
We present our results and discussions in Sec. \ref{results}.
Finally, a conclusion is given.




