\documentclass{article}
\usepackage{amsmath} 
\usepackage{graphicx} 
\usepackage{authblk}
\usepackage{url}
\usepackage{hyperref}
\usepackage{multirow}
\usepackage{amssymb}
\usepackage[linesnumbered,boxed]{algorithm2e}

\topmargin 0.0cm
\oddsidemargin 0.2cm
\textwidth 16cm 
\textheight 21cm
\footskip 1.0cm

\title{Quantum Monte Carlo method and and Its Application in Physics}
\author{Mengzhi Chen}
\affil{Department of Physics and Astronomy, Michigan State University}
\date{}

\begin{document}
	\maketitle
	\begin{abstract}\label{abstract}
Monte Carlo (MC) method is widely used in scientific computations.
This method relies on huge amount of random samplings to obtain numerical results.
MC method shows high efficiency and precision when dealing with many complex problems, such as non-analytic and high demential integrals. 
In this report we apply MC method together with variational principle to calculate the ground state energy of two electrons moving in a harmonic oscillator(H.O.) potential.
We assumed two forms of trial wave functions with parameters. 
By using MC method, we find their optimistic parameters corresponding to ground state energies.
The agreements with results from diagonalization method (in Project 2) validate our MC simulations.
We also test the virial theorem for different relative strength between H.O. potential and Coulomb repulsion.

	\end{abstract}

\section{Introduction}\label{intro} 
\input{introduction}

\section{A physicical system}\label{phys}
\input{physics}
	
	
\section{Numerical methods}\label{method}
	\subsection{Metropolis algorithm}\label{metro}
	\input{metro}

	
	\subsection{Variational Monte Carlo(VMC) method} \label{vmc}
	\input{vmc}


	
\section{Results and discussions}\label{results}

	\subsection{First trail wave function}
	\input{first}


	\subsection{Second trail wave function}
	\input{second}

	
	\subsection{Test of virial theorem}
	\input{virial}

	
\section{Conclusions}\label{conclude}
	\input{con}
	
	\section*{Acknowledgments}
	We are grateful for the sincere guidance from Prof. Morten Hjorth-Jensen. 
	
	\nocite{*} 
	\bibliographystyle{unsrt}
	\bibliography{proj4_ref}
\end{document}