Suppose a first-order ordinary differential equation

\begin{equation}
	\frac{dx}{dt} = f(x,t);\ t\in[t_0,t_1]
\end{equation}
with a initial value $x_0$ at time $t_0$. 

In order to solve this problem, we first discretize the region $ [t_0,t_1]$ into N subintervals with a step $h$, so we get the relation

\begin{equation}
	h=\frac{t-t_0}{N}.
\end{equation}
Then, We have discretized $x_i=x(t_i=t_0+ih)$ where $i$ is a integer between 0 and $N$.

Using Taylor expansion, we get 

\begin{equation}
	\label{eulerexpan}
	x_{i+1} = x_i + \frac{dx_i}{dt}h + \frac{d^2x_i}{dt^2}h^2 + O(h^3),
\end{equation}
where i goes from 0 to $N-1$.
The Euler's forward method truncates at the second term of the above equation.
Thus, Eq. \ref{eulerexpan} becomes

\begin{equation}
	\begin{aligned}
		x_{i+1} & = x_i + \frac{dx_i}{dt}h + O(h^2) \\ 
				& = x_i + f(x_i,t_i)h + O(h^2).
	\end{aligned}
\end{equation}
We can see it's a one-step method with a local error $O(h^2)$.

Getting back to the Earth-Sun system, we can formulate Eq. \ref{earthsunodes} to

\begin{equation}
	\left\{  
             \begin{array}{lr}  
             	x_{i+1} = x_i + v_x^{i}h \\
				y_{i+1} = y_i + v_y^{i}h \\
             	v_x^{i+1} = v_x^{i} - \frac{4\pi^2x_i}{r_i^3}h\\
             	v_y^{i+1} = v_y^{i} - \frac{4\pi^2y_i}{r_i^3}h.
			\end{array}  
	\right.	
\end{equation} 
in unit of AU for length, year (yr) for time and $M_{\mathrm{S}}$ for mass. 
We will keep using this unit system in our report and calculations. 
Starting from some initial conditions, we can simply solve out time evolution of the Earth iteratively. 
Our implementation of this method with a  circular orbit initial conditions is shown in Algorithm \ref{alg::euler}. 

\begin{algorithm}[tb]
	\caption{The Euler's forward method for the Earth-Sun system. It initials from a circular orbit.}
	\label{alg::euler}
	\KwIn{$x_0=1$, $y_0=0$, $v_{x}^0=0$, $v_{y}^0=2\pi$, $M_{\mathrm{E}}$, $M_{\mathrm{S}}$, $G$ }
	\KwOut{$\vec{x}$=($x_0$,$x_1$,...,$x_N$), $\vec{y}$, $\vec{v}_x$, $\vec{v}_y$, $\vec{E}_k$, $\vec{E}_p$, $\vec{E}$, $\vec{L}_z$} 
	$r$ = sqrt($x_0^2$+$y_0^2$)\;
	// $i$ is deferent time points $t_i=t_0+ih$\;
    \For{$i=1;i<=N;i++$}
    { $x_{i} = x_{i-1} + v_x^{{i-1}}h$;
    $y_{i} = y_{i-1} + v_y^{{i-1}}h$\;
    $v_x^{i} = v_x^{i-1} - \frac{4\pi^2x_{i-1}}{r_{i-1}^3}h$;
    $v_y^{i} = v_y^{i-1} - \frac{4\pi^2y_{i-1}}{r_{i-1}^3}h$\;
    $r$ = sqrt($x_i^2$+$y_i^2$)\;
    // Kinetic energy $E_k$, Potential energy $E_k$, Total energy $E$, Angular momentum in $\hat{z}$ $L_z$\;
    $E_k^i$ = 0.5$M_{\mathrm{E}}$($(v_x^{i})^2$+$(v_y^{i})^2$);
    $E_p^i$ = $-\frac{GM_{\mathrm{E}}M_{\mathrm{S}}}{r}$\;
    $E^i$ = $E_p^i$ + $E_k^i$;
    $L_z^i$ = $M_{\mathrm{E}}(x_iv_y^{i}-y_iv_x^{i})$\;
    }
	return $\vec{x}$, $\vec{y}$, $\vec{v}_x$, $\vec{v}_y$, $\vec{E}_k$, $\vec{E}_p$, $\vec{E}$, $\vec{L}_z$\;
\end{algorithm}

We can see that the Euler's forward method is easy to be realized.
 However, it has vital defects that it violates the energy conservation and time reversibility.
 The total energy increases with time in the Euler's forward method.
 That's why we need the Velocity-Verlet method to describe physical systems.
